---
title: "Process Mohammed Data"
author: "Stephanie Hickey"
date: "2019-11-07"
output: html_document
---

## Data acquisition
### 11-07-19

Data from [_Single-Cell Landscape of Transcriptional Heterogeneity and Cell Fate Decisions during Mouse Early Gastrulation, Mohammed et al. 2017_](http://www.sciencedirect.com/science/article/pii/S2211124717309610)

#### The filtered counts data was downloaded from GEO: [GSE100597](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE100597)

#### Make a meta data table and clean up the rows and columns

```{r, eval=FALSE}
library(tidyr)
library(stringr)
library(base)

setwd("/mnt/home/hickeys6/Documents/Ralston/projects/Mohammed2017/data")
dat = read.delim("GSE100597_count_table_QC_filtered.txt")
rownames(dat) = dat$X
dat$X = NULL
```

In **dat** rows are genes and columns are cells. I want to make a metadata table from the information in the columns, but the formatting is inconsistant. 
For example:

*  Sometimes it's "cell"        
*  Sometimes it's "Cell"    
*  And Sometimes it's "Cell_"   

```{r, eval=FALSE}
meta = colnames(dat) 
meta = str_replace(meta, "cell", "Cell")
meta = str_replace(meta, "Cell_", "Cell")
meta = str_replace(meta, "embryo", "Embryo")
meta = str_replace(meta, "single", "Single")
meta = str_replace(meta, "double", "Double")
colnames(dat) = meta
```

Now I'll separate cell names to get metadata

*  Some cells have "P" info, some don't. I don't know what the "P" info is for
*  None the less, I don't want to lose information

```{r, eval=FALSE}
metadf = as.data.frame(meta)
colnames(metadf) = "CellID"
metadf = separate(metadf, CellID, c("Stage","Rest"), sep = "_", extra = "merge", remove = FALSE)
# make a df with only cells with "P" info
metaP = metadf[grep("P", metadf$Rest),]
# make a df with only cells without "P" info
meta_noP = metadf[grep("^[^P]", metadf$Rest),]
# separate cell names to get metadata
metaP = separate(metaP, Rest, c("P","Cell_Number", "Embryo", "Singlet_v_Doublet"), sep = "_")
meta_noP = separate(meta_noP, Rest, c("Cell_Number", "Embryo", "Singlet_v_Doublet"), sep = "_")

# add a "P" column to noP and order columns the same as metaP
meta_noP$P = rep("NA", nrow(meta_noP))
meta_noP = meta_noP[,c(1,2,6,3,4,5)]

# rbind the meta data
meta_final = rbind(metaP, meta_noP)
```

The meta data table is made, but the rownames are the gene symbol plus a numeric code I don't recognize.

```{r, eval=FALSE}
# remove the numbers and keep the gene symbols
gene_codes = rownames(dat)
gene_symbols = str_extract(gene_codes, "[^_]+")
head(gene_symbols)
```

I want to make the gene symbols the row names. But I need to check to see if any of the symbols are duplicated:

I'll remove rows with duplicated rownames and save the data
```{r, eval=FALSE}
# remove the duplicated symbols
symbols_no_dup = gene_symbols[!duplicated(gene_symbols)]
gene_codes_no_dup = gene_codes[!duplicated(gene_symbols)]
dat_no_dup = dat[gene_codes_no_dup,]
rownames(dat_no_dup) = symbols_no_dup
rawCounts = dat_no_dup

save(rawCounts,meta_final, file = "mohammed_clean.Rdata")
```

## Clustering with Seurat
### 12-11-19

```{bash, eval=FALSE}
ssh dev-intel16-k80.i
srun -N 1 -t 5:00:00 --mem 64G --pty bash
module purge
module load GCC/7.3.0-2.30 OpenMPI/3.1.1
module load R/3.5.1-X11-20180604
R --vanilla
```

#### Make basic QC-plots and cluster
```{r, eval=FALSE}
########################################################################
setwd("/mnt/home/hickeys6/Documents/Ralston/projects/Mohammed2017/data")
library(Seurat)
library(dplyr)
library(ggplot2)
########################################################################

load("mohammed_clean.Rdata")
ls()
dat = CreateSeuratObject(rawCounts)
# check to see if there are mito genes in the data
mito.genes <- grep(pattern = "^mt-", x = rownames(rawCounts), value = TRUE, ignore.case = TRUE)
mito.genes
# none to be found...

# log normalize the data
dat <- NormalizeData(dat)

# find variable genes
dat <- FindVariableFeatures(dat, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(dat), 10)

# scale the data
all.genes <- rownames(dat)
dat <- ScaleData(dat, features = all.genes)

# run PCA
dat <- RunPCA(dat, features = VariableFeatures(object = dat))

# determine dimentionality
dat <- JackStraw(dat, num.replicate = 100)
dat <- ScoreJackStraw(dat, dims = 1:20)

# cluster the cells
# jackstraw suggests the first 10 PCs
dat <- FindNeighbors(dat, dims = 1:10)
dat <- FindClusters(dat, resolution = 0.5)

save(dat, file = "Mohammed2017_seurat.rdata")
```
